---
title: "GOOG_Model_Instruction_Manual"
author: "Christian DeJong"
date: "2023-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Alphabet Inc. Stock Price Model Instruction Manual

## 1. Gather Data from their respective website and copy the tables from 2004-2022 into separate excel files:

### - GOOGL went public in 2004 so no public data exists before 2004

-   GOOGL Stock Price : <https://www.macrotrends.net/stocks/charts/GOOGL/alphabet/stock-price-history>
-   Inflation Rate : <https://www.usinflationcalculator.com/inflation/current-inflation-rates/>
-   Financial Statement Data (Operating Cash Flow and Revenues) : <https://www.sec.gov/Archives/edgar/data/1652044/000165204422000019/goog-20211231.htm>
-   Unemployment Rate : <https://www.bls.gov/charts/employment-situation/civilian-unemployment-rate.htm>
-   Market Return (S&P 500 index) : <https://www.macrotrends.net/2324/sp-500-historical-chart-data>

## 2. For ease of use, save a copy of each excel file in a .csv format and upload the .csv file into a GitHub repository.

## 3. Load Libraries

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## 4. Load Data into temporary variables using read_csv()

```{r}
EmployRate <- read_csv("https://raw.githubusercontent.com/cdejong1/DATA316/main/Stock%20Model%20Data/EmploymentRate.csv")
FinStatement <- read_csv("https://raw.githubusercontent.com/cdejong1/DATA316/main/Stock%20Model%20Data/GOOG%20Financial%20Statment%20Data.csv")
StockPrice <- read_csv("https://raw.githubusercontent.com/cdejong1/DATA316/main/Stock%20Model%20Data/GOOG%20stock%20price.csv")
InflationRate <- read_csv("https://raw.githubusercontent.com/cdejong1/DATA316/main/Stock%20Model%20Data/InflationRate.csv")
MktReturn <- read_csv("https://raw.githubusercontent.com/cdejong1/DATA316/main/Stock%20Model%20Data/MktReturn.csv")
```

## 5. Combine data frames using the merge() function using Date as the primary key

```{r}
df1 <- merge(EmployRate, FinStatement, by = "Date")
df2 <- merge(df1, StockPrice, by = "Date")
df3 <- merge(df2, InflationRate, by = "Date")
GOOG <- merge(df3, MktReturn, by = "Date")
```

## 6. Format Date Column for ease of use when plotting

```{r}
GOOG$Date <- as.Date(GOOG$Date, format = "%m/%d/%y")
```

## 7. Create a new column for change in stock price

### - First using arrange(Date) to arrange the columns by date, then use mutate() to create the new column as a function of (StockPrice - lag(StockPrice)). Default for StockPrice is year 1

```{r}
GOOG <- GOOG %>%
  arrange(Date) %>%
  mutate(chng_in_price = StockPrice - lag(StockPrice, default = first(StockPrice)))
```

## 8. Using ggplot() and geom_line() create a plot to visualize the Stock Price over time to get a feel for the data we are modeling.

### - Use Date as the x-axis and StockPrice for the y-axis.

```{r}
ggplot(GOOG, aes(x= Date, y = StockPrice)) +
  geom_line()
```

## 9. Create a linear regression model using lm()

### - Use chng_in_price as the response variable, and MktReturn, Revenue, OCF, EmployRate and InflationRate as explanatory variables. Using data = yourDataFrame.

### -View your model estimates using the summary() function on your model.

```{r}
mod.4 <- lm(chng_in_price ~ MktReturn + Revenue + OCF + EmployRate + InflationRate, GOOG)
summary(mod.4)
```
